%===========================
% Part II Progress Report
%===========================
\documentclass[11pt,a4paper]{scrartcl}

%---- Encoding & typography
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{microtype}
\usepackage{newpxtext,newpxmath}
\linespread{1.04}

%---- Layout
\usepackage[margin=1in]{geometry}
\usepackage{graphicx}
\KOMAoptions{parskip=half}

%---- Colors, links, and headings
\usepackage[dvipsnames]{xcolor}
\definecolor{Accent}{HTML}{4F46E5}
\definecolor{AccentDark}{HTML}{3730A3}
\usepackage{hyperref}
\hypersetup{
  colorlinks=true,
  linkcolor=AccentDark,
  urlcolor=Accent,
  citecolor=Accent
}

%---- Section styling
\addtokomafont{section}{\large\bfseries\color{AccentDark}}
\addtokomafont{subsection}{\normalsize\bfseries\color{AccentDark}}
\RedeclareSectionCommand[beforeskip=10pt,afterskip=4pt]{section}
\RedeclareSectionCommand[beforeskip=8pt,afterskip=3pt]{subsection}

%---- Lists
\usepackage{enumitem}
\setlist[itemize]{itemsep=0.1em, topsep=0.2em}

%---- Callout boxes
\usepackage[most]{tcolorbox}
\tcbset{
  colback=Accent!4,
  colframe=Accent!60!black,
  boxrule=0.6pt,
  arc=2mm,
  left=6pt,right=6pt,top=5pt,bottom=5pt
}
\newtcolorbox{successbox}{title=\textbf{Success Criteria Status}}

%---- Header / footer
\usepackage{fancyhdr}
\pagestyle{fancy}
\fancyhf{}
\lhead{\textit{Part II Progress Report}}
\rhead{\thepage}
\renewcommand{\headrulewidth}{0.4pt}

%---- Title
\title{\vspace{-1.5em}\textbf{Part II Progress Report}}
\subtitle{A Workbench for Comparing Voice-LLM Architectures}
\author{
  \textbf{Vedaangh Rungta} (vr373@cam.ac.uk)\\[0.2em]
  \small Supervisors: Dr Soumya Banerjee, Prof Nic Lane \quad DoS: Janet Davis
}
\date{5 February 2026}

\begin{document}
\maketitle
\vspace{-1.5em}
\hrule height 0.6pt \vspace{0.6em}

\begin{successbox}
\textbf{Criterion 1} (Workbench): An API that can be hosted remotely to interact with pre-trained models through a unified interface and evaluate their performance using VoiceBench. \textbf{Status: Not yet started.}\\[0.3em]
\textbf{Criterion 2} (Llama-Omni): A working Llama-Omni re-implementation interoperable with Qwen3-4B and Qwen3-30B, accessible through the API. \textbf{Status: Partially complete.} Qwen3-4B trained; Qwen3-30B and API integration pending.
\end{successbox}

\section{Work Completed}

The InstructS2S-200K dataset's speech tokens were incompatible with our K=1000 HuBERT vocabulary, so I built a pipeline using Azure TTS to re-synthesize 423K samples and extract discrete units via mHuBERT and K-means clustering, producing 422,790 usable samples.

I implemented the encoder path with Whisper-large-v3 and a 2-layer MLP adapter that downsamples by 5$\times$ and projects into Qwen3-4B's embedding space. Training took 7.5 hours on a single H100 for 3 epochs, achieving a final validation loss of 1.60.

The decoder is a non-autoregressive transformer (12 layers, 1024 hidden dimension) trained with CTC loss to output discrete units for a HiFi-GAN vocoder. Training completed in 16.5 hours with a final validation loss of 0.74.

The end-to-end inference pipeline is operational. Output speech is coherent for 7--8 seconds before degrading, likely because our decoder is smaller than the original paper's (1024 vs 4096 hidden dimensions).

\section{Difficulties}

Dataset incompatibility added one week of unplanned work. GPU availability on AWS was inconsistent, with H100 instances frequently unavailable. Competing coursework deadlines also consumed time during Michaelmas, and a period of sickness over the break led to a backlog forming.

\section{Schedule and Recovery}

The original schedule targeted Qwen3-30B training during Michaelmas vacation and a first draft by Lent week 2 to allow time for submitting complementary work to workshops. I am now submitting separate research to those workshops instead. The revised plan is:

\subsection*{Lent term (Revised)}
\begin{itemize}
  \item \textbf{Week 3:} Decoder ablations to improve output consistency.
  \item \textbf{Week 4:} Qwen3-30B training run.
  \item \textbf{Week 5:} API development and VoiceBench evaluation.
  \item \textbf{Weeks 6--8:} Extensions.
\end{itemize}

\subsection*{Lent break / Easter}
\begin{itemize}
  \item \textbf{Lent break weeks 1--2:} Dissertation writing.
  \item \textbf{Lent break weeks 3--4 / Easter:} Buffer.
\end{itemize}

I have also secured increased GPU quotas on both AWS and GCP, providing access to multiple H100s, A100s, and B200s, which resolves the previous availability issues.

\end{document}
